---
title: "Online News Analysis - Predicting Media Shares by Article Characteristics"
author: "Matt Kasle"
date: "10/15/2020"
output: rmarkdown::github_document
params:
  day_of_week: "monday"
  toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# `r paste(params$day_of_week, "articles")`

## Introduction 
This dataset summarizes a heterogeneous set of features about articles published by Mashable in a period of two years. The goal is to predict the number of shares in social networks (popularity).

There are 61 attributes, 58 which are predictive attributes, 2 of which are non-predictive (url and timestamp), and 1 that is the target.

The types of veriables include the number of words in the article, the number of words in the title, the positivity and sentimentality of the article, the article's subject, the number of keywords used, and much more.

The objective of this project is to predict the number of social media shares using two different tree-based algorithms. The first algorithm will be a non-ensemble regression tree, and the second algorithm will be a boosted trees algorithm, which is a state-of-the-art classification technique.

The other objective of this project is to create an automated report that outputs the classification analysis for each articles published on each weekday. 

The required packages to run this analysis are tidyverse, caret, tree, and patchwork.

## Data

First, we need to read in the data and set up the subsetting of the data into a single weekday. Then, we'll split the data set to a training and test set for training and evaluating the classificaiton models. A 70/30 train-test split will be used.
```{r warning=FALSE, message = FALSE}
set.seed(1)
library(tidyverse)
library(caret)
library(tree)

news <- read_csv("OnlineNewsPopularity.csv")
```

```{r, message=FALSE}
varDayOfWeek <- parse(text=paste0("weekday_is_", params$day_of_week))
news <- news %>% filter(eval(varDayOfWeek) == 1)

news <- news %>% select(-url, -timedelta, -weekday_is_monday, -weekday_is_tuesday, -weekday_is_wednesday,
                        -weekday_is_thursday,-weekday_is_friday, 
                        -weekday_is_saturday, -weekday_is_sunday, -is_weekend)

newsIndex <- createDataPartition(news$shares, p = 0.3, list = FALSE)
newsTrain <- news[newsIndex, ]
newsTest <- news[-newsIndex, ]
```

## Summarizations
Below are summary statistics of the online news data set, such as the number of observations in the train set, a numerical summary of the response variable (number of social media shares), a distribution of the response, and relationships between the response and interesting variables in the dataset.

Number of rows in training set:
```{r}
nrow(newsTrain)
```

Summary of response variable:
```{r}
summary(newsTrain$shares)
```
It is important to make note of the minimum and maximum of the response variable, as well as the median and quartiles. For future analysis, it may be best to remove outliers. 

Distribution of response variable:
```{r, message=FALSE}
ggplot(data = newsTrain, aes(x = shares)) +
  geom_histogram() + xlab("Shares") + ggtitle(paste("Distribution of Shares in Training Data - ",
                                                    params$day_of_week))
```

Below is a plot of four interesting variables to the response variable, shares. We want to see if there are any intersting patterns here.
```{r, message=FALSE}
library(patchwork)
par(mfrow=c(2,2))
plot1 <- ggplot(data=news, aes(x = news$n_unique_tokens, y = shares)) + geom_point(stat = "identity") +
    labs(x="Unique Words in Article", y="Shares")

plot2 <- ggplot(data=news, aes(x = news$rate_positive_words, y = shares)) + geom_point(stat = "identity") +
    labs(x="Rate of Positive Words", y="")

plot3 <- ggplot(data=news, aes(x = news$rate_negative_words, y = shares)) + geom_point(stat = "identity") +
    labs(x="Rate of Negative Words", y="Shares")

plot4 <- ggplot(data=news, aes(x = news$global_sentiment_polarity, y = shares)) + geom_point(stat = "identity") +
    labs(x="Sentiment Polarity", y="")

plot1 + plot2 + plot3 + plot4
```

There are six different article subjects. It would be interesting to know if some subjects are more popular than others, so we can plot the median shares by the different article types.
```{r}
news['article_type'] <- ifelse(news$data_channel_is_lifestyle == 1, "Lifestyle",
       ifelse(news$data_channel_is_entertainment == 1, "Entertainment",
       ifelse(news$data_channel_is_world == 1, "World",
       ifelse(news$data_channel_is_bus == 1, "Bus",
       ifelse(news$data_channel_is_socmed == 1, "SocMed",
       ifelse(news$data_channel_is_tech == 1, "Tech","Other"))))))

ggplot(data=news, aes(x = article_type, y = shares)) + geom_bar(stat = "summary", fun.y = "median") +
    labs(x="Article Type", y="Median Shares", title="Median Shares by Article Type")

# drop column used for plotting
news <- news %>% select(-article_type)
```


## Modeling

There should be text describing the type of model you are fitting, your fitting process, and the final chosen model (this last part is to be automated so I don’t expect you to explicitly interpret that model, but you should be able to display something about the final model chosen on the training data).

### Regression Tree Model

The first model fit to the data will be a regression tree. We'll use leave-one-out cross-validation to determine the optimal size of the model, as defined by number of splits. By plotting the fitted tree, we can see the deviance by tree size (larger deviance means a better fit).

```{r}
treeFit <- tree(shares ~ ., data = newsTrain)
summary(treeFit)
pruneFit <- cv.tree(treeFit,
                  K=nrow(newsTrain)-1
                  )


pruneFitFinal <- prune.tree(treeFit, best = pruneFit$size[1]) 

plot(pruneFit$size ,pruneFit$dev ,type="b")
```

### Boosted tree model
Next, we'll fit a boosted regression tree. The boosted tree algorithm has a few hyperparameters, and we'll use repeated 10-fold cross-validation to determine the optimal hyperparameter values. The hyperparameters of the optimal boosted tree is printed below, as well a summary of each fitted model.
```{r, message=FALSE}
fitControl <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           ## repeated ten times
                           repeats = 5
                           )

boostedFit <- train(shares ~ ., data = newsTrain, 
                 method = "gbm", 
                 trControl = fitControl, 
                 verbose = FALSE
                 )

boostedFit$bestTune
```

```{r, message=FALSE}
boostedFit$results
```


### Linear regression model
Next, we'll fit a siple multiple linear regression model on train data. Summary of the model is printed below.
```{r}
linearfit<-lm(shares~.,data=newsTrain)
summary(linearfit)
```


## Model Evaluations
Finally, we'll evaluate the performance of each model by seeking the lowest root mean squared error of its predictions on the test dataset when compared to the actual values in the dataset. This should be a good approximation of the model's performance on unseen data.

### Regression Tree

Below is the RMSE of the optimal (non-ensemble) regression tree:
```{r, message=FALSE}
treePred <- predict(pruneFitFinal, newdata = dplyr::select(newsTest, -shares))
sqrt(mean((treePred-newsTest$shares)^2))
```

### Boosted Trees

Below is the RMSE of the optimal boosted regression tree:
```{r, message=FALSE}
boostedTreePred <- predict(boostedFit, newdata = dplyr::select(newsTest, -shares))
sqrt(mean((boostedTreePred-newsTest$shares)^2))
```


### linear regression

Below is the RMSE of the multiple linear model:
```{r, message=FALSE,warning=FALSE}
linearPred<-predict(linearfit, newdata=dplyr::select(newsTest, -shares))
sqrt(mean((linearPred-newsTest$shares)^2))
```
